---
title: "P85105_hw3_jw3597"
output: github_document
---

#### Name: Jessie Wong 
#### Date: October 7, 2018

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```

#### Problem 1

This problem uses the BRFSS data. The data is loaded from the p8105.datasets package.  

First we will do some data cleaning: 

* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels from “Excellent” to “Poor”

```{r}
library(p8105.datasets)
data("brfss_smart2010")

# I am formatting the data using appropriate variable names and then only using responses from "Excellent" to "Poor":
brfss_smart2010 = janitor::clean_names(brfss_smart2010) %>%
  rename(state = locationabbr, county = locationdesc, health_issue = class, health_issue_topic = topic, health_question = question, health_response = response, lower_95CI = confidence_limit_low, upper_95CI = confidence_limit_high) %>% 
  filter(health_response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
           mutate(health_response = factor(health_response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>%
  arrange(health_response)

```

In 2002, the states observed at 7 locations are

#### Problem 2

This problem uses the Instacart data. The data is loaded from the p8105.datasets package (called instacart).  

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each): 

* How many aisles are there, and which aisles are the most items ordered from?
* Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
* Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
data("instacart")
```

#### Problem 3

This problem uses the NY NOAA data. The data is loaded from the p8105.datasets package.  

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

* Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
* Make a two-panel plot showing the average temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
* Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
data("ny_noaa")
```





