---
title: "P85105_hw3_jw3597"
name: "Jessie Wong"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
```

#### Problem 1

This problem uses the BRFSS data. The data is loaded from the p8105.datasets package.  

First we will do some data cleaning: 

* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels from “Excellent” to “Poor”

```{r}
library(p8105.datasets)
data("brfss_smart2010")

# I am formatting the data using appropriate variable names, including only using responses from "Excellent" to "Poor", and then organizing the responses as a factor:
brfss_smart2010 = janitor::clean_names(brfss_smart2010) %>%
  rename(state = locationabbr, county = locationdesc, health_issue = class, health_issue_topic = topic, health_question = question, health_response = response, lower_95CI = confidence_limit_low, upper_95CI = confidence_limit_high) %>% 
  filter(health_response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
           mutate(health_response = factor(health_response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))) %>%
  arrange(health_response)
```

Using this dataset, I will answer the following (commenting on the results of each):

* In 2002, which states were observed at 7 locations?
```{r}
brfss_2002 = brfss_smart2010 %>% 
  filter(year == "2002") %>%
  group_by(state) %>%
  summarize(n = n_distinct(county)) %>%
  filter(n == 7)
```
In 2002, the states observed at 7 locations are CT, FL, and NC.

* Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.
```{r}
gg2002_2010 = brfss_smart2010 %>%
  group_by(state, year) %>%
  distinct(county) %>%
  count(year,state)

ggplot(gg2002_2010, aes(x = year, y = n, color = state)) + geom_line(se = FALSE)
```

* Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
```{r}
table1 = brfss_smart2010 %>%
  filter(state == "NY", health_response == "Excellent", year %in% c("2002", "2006", "2010")) %>%
  spread(key = health_response, value = data_value) %>%
  group_by(year) %>%
  summarize(mean_excellent = mean(Excellent/100), sd_excellent = sd(Excellent/100))
```

* For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.
```{r}
panel_plot = brfss_smart2010 %>%
  group_by(state, year, health_response) %>%
  summarize(average_proportion = mean(data_value/100))

ggplot(panel_plot, aes(x = year, y = average_proportion, color = state)) + 
  geom_line(alpha = 0.5) + 
  facet_grid(~health_response)
  
```

#### Problem 2

This problem uses the Instacart data. The data is loaded from the p8105.datasets package (called instacart).  

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations. Then, do or answer the following (commenting on the results of each): 

* How many aisles are there, and which aisles are the most items ordered from?
* Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
* Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
data("instacart")

# Identifying the number of aisles and aisles with the most ordered items
uni = distinct(instacart, aisle_id)
count(uni)
  
sort(table(instacart$aisle), decreasing = TRUE) %>%
  head()
```
"Instacart" is a dataset that shows the online grocery orders of its users, containing `r nrow(instacart)` rows and `r ncol(instacart)` columns. Some of the key variables are the product name, aisle name and number where the product is found/categorized in, whether the item(s) were reordered, and the order ID that indicates the specific order placed by a customer.

* There are `r count(uni)` aisles and the aisles with the most ordered items are aisle 83 "fresh vegetables", 24 "fresh fruits", and 123 "packaged vegetable fruits".

```{r}
# Plot that shows the number of items ordered in each aisle. Aisles will be ordered sensibly and the plot will be organized to be readable to others
items = instacart %>%
  group_by(aisle, aisle_id) %>% 
  count(aisle) %>% 
  rename(items_ordered = n) %>%
  arrange(aisle)

ggplot(items, aes(x = aisle_id, y = items_ordered)) + 
  geom_point(alpha = .5)
```

```{r}
# Table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits
popular = instacart %>%
  group_by(aisle) %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  summarize(most_popular = max(product_name))
```
* The most popular item in "baking ingredients" aisle is the zero calorie sweeter, the most popular item in "dog food care" is Zero Grain Turkey Dog Food, and the most popular item in "packaged vetables fruits" is Zucchini Squash, Baby Courgette.

```{r}
# Table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)
  
  
```


#### Problem 3

This problem uses the NY NOAA data. The data is loaded from the p8105.datasets package.  

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):
```{r}
data("ny_noaa")
sum(is.na(ny_noaa))
```
"Ny_noaa" is a dataset that contains information about weather across different countries and terrorities. There are `r ncol(ny_noaa)` total and `r nrow(ny_noaa)` rows of entries. Among the 7 variables in the dataset, a few key ones are "ID", which specifies the specific weather station, where the data in the row corresponds to and collected from; "prcp" is the precipation amount that is measured in tenth of mm units; "snow" is the amount of snowfall that was measured in mm units; "date" is an important variable because it indicates the date of observation when the data in the row was recorded. Missing data is a major issue because there are `r sum(is.na(ny_noaa))` missing information data entries across the entire dataset and this could cause incomplete or inaccurate analysis of the weather across the stations.

* Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
```{r}
ny_noaa = janitor::clean_names(ny_noaa) %>%
  separate(date, into = c("year", "month", "day"), sep = "-")
  
```

* Make a two-panel plot showing the average temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}

```

* Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
```{r}

```
